Causal Impact of Masks, Policies, Behavior on Early Covid-19 Pandemic in the U.S.

### 2020-09-08 tex/covidEffectOfNPI-revision.tex
info  update

The specification of the model above grasps one-period responses.  The dynamics over multiple periods will be induced by the evolution
of information variables, which include time, lagged and integrated values of outcome:\footnote{Our empirical analysis also considers a specification in which information variables include lagged national cases/deaths as well as lagged behavior variables.}
\begin{equation}\label{eq:I}\tag{I}
 {\icolor I_{it} } := I(t, {\ycolor Y_{it}},  { \icolor  I_{i, t-1} })' := \Big (g(t) , {\ycolor Y_{it}},  \sum_{m=1}^{ \lfloor t/\ell \rfloor}
{\ycolor Y_{i,t - \ell m}} \Big )  \end{equation}
 for  each  $t \in \{0,1,...,T\}$, where $g$ is deterministic function of time, e.g., month indicators,
 assuming that the log of new cases at time $ - t \ \mathrm{mod} \ \ell $ is zero, for notational convenience.
 \footnote{The general formula  for  $I_{i, t-1}$ is $
S_{i, t, \ell } + \sum_{m=1}^{ \lfloor t/\ell \rfloor} Y_{i,t - \ell m}$, where $ S_{i, t, \ell }$ is the initial condition, the log of new cases at time $- t \ \mathrm{ mod } \ \ell$.}

In this structure, people respond to both global information, captured by a function of time such as month dummies, and local information sources, captured by the local growth rate and the total number of cases. The local information also captures the persistence of the growth rate process.  We model the reaction of people's behavior via the term ${\icolor \gamma'I_{t}}$ in the behavior equation.  The dynamical system requires initialization, for example we can set  ${\ycolor Y_{it}} = 0$ for $t \leq 0$. The  lagged values of behavior variable may be also included in the information set, but we postpone this discussion after the main empirical results are presented.

@@

\begin{quote}
\textsc{Structural Equations Model (SEM)}.  Let $i \in \{1,..., N\}$ denote the observational unit,  $t$ be the time periods, and $\ell$ be the time delay.  (1) For each $i$ and $t \leq -\ell$, the confounder, information, behavior, and policy variables $W_{it}, I_{it},  B_{it}, P_{it}$ are determined outside of the model, and the outcome variable $Y_{i, t+\ell}$ is determined by factors outside of the model for $t \leq 0$.  (2) For each $i$ and $t \geq -\ell$, confounders $W_{it}$ are  determined by factors outside of the mode, and information variables $I_{it}$ are determined  by (I); policy variables $P_{it}$ are determined by setting $\iota = I_{it}$  in (P) with  a realized stochastic shock $\varepsilon^p_{it}$ that obeys the exogeneity condition (E);  behavior variables $B_{it}$ are determined by setting $\iota = I_{it}$ and $p= P_{it}$ in  (SO) with a shock $\varepsilon^b_{it}$ that obeys (E); finally, the outcome $Y_{i, t + \ell}$ is realized by setting $\iota = I_{it}$, $p= P_{it}$, and $b = B_{it}$ in  (SO) with a shock $\varepsilon^y_{it}$ that obeys (E). \end{quote}

@@

Given the policies, we generate the counterfactual
outcomes, behavior, and information by propagating the dynamic equations:
\begin{equation} \label{eqn:CF-SEM} \tag{CF-SEM}
  \begin{aligned}
& {\ycolor Y^\star_{i,t+\ell}}  & := & Y_{i,t+\ell} ( {\bcolor B^\star_{it} } , {\pcolor P^\star_{it}}, {\icolor I^\star_{it}}), \\
& {\bcolor B^\star_{it} } & := &   B_{it}({\pcolor P^\star_{it} } ,{\icolor I^\star_{it}}), \\
& {\icolor I^\star_{it} }& := &  I ({ t,  {\ycolor Y^\star_{it}}, \icolor I^\star_{i,t-\ell}}),  \end{aligned}
\end{equation}
with the same initialization as the factual system up to $t \leq 0$. In stating this counterfactual system of equations, we assume that structural outcome equations (SO) and information equations (I) remain invariant and so do the stochastic shocks, decomposed
into observable and unobservable parts. Formally, we record this assumption and above discussion as follows.

>>>>

The specification of the model above grasps one-period responses.  The dynamics over multiple periods will be induced by the evolution
of information variables, which include time, lagged and integrated values of outcome:\footnote{Our empirical analysis also considers a specification in which information variables include lagged national cases/deaths as well as lagged behavior variables.}
\begin{equation}\label{eq:I}\tag{I}
 {\icolor I_{it} } := I_t( {\ycolor Y_{it}},  { \icolor  I_{i, t-\ell} }) := \Big (g(t) , {\ycolor Y_{it}},  \sum_{m=0}^{ \lfloor t/\ell \rfloor}
{\ycolor Y_{i,t - \ell m}} \Big )'  \end{equation}
 for  each  $t \in \{0,1,...,T\}$, where $g$ is deterministic function of time, e.g., month indicators,
 assuming that the log of new cases at time $ - t \ \mathrm{mod} \ \ell $ is zero, for notational convenience.
 \footnote{The general formula  for  $I_{i, t-1}$ is $
S_{i, t, \ell } + \sum_{m=1}^{ \lfloor t/\ell \rfloor} Y_{i,t - \ell m}$, where $ S_{i, t, \ell }$ is the initial condition, the log of new cases at time $- t \ \mathrm{ mod } \ \ell$.} In this structure, people respond to both global information, captured by a function of time such as month dummies, and local information sources, captured by the local growth rate and the total number of cases. The local information also captures the persistence of the growth rate process.  We model the reaction of people's behavior via the term ${\icolor \gamma'I_{t}}$ in the behavior equation.  The  lagged values of behavior variable may be also included in the information set, but we postpone this discussion after the main empirical results are presented.

@@

\begin{quote}
\textsc{Structural Equations Model (SEM)}.  Let $i \in \{1,..., N\}$ denote the observational unit,  $t$ be the time periods, and $\ell$ be the time delay.  (1) For each $i$ and $t \leq -\ell$, the confounder, information, behavior, and policy variables $W_{it}, I_{it},  B_{it}, P_{it}$ are determined outside of the model, and the outcome variable $Y_{i, t+\ell}$ is determined by factors outside of the model for $t \leq 0$.  (2) For each $i$ and $t \geq -\ell$, confounders $W_{it}$ are  determined by factors outside of the model, and information variables $I_{it}$ are determined  by (I); policy variables $P_{it}$ are determined by setting $\iota = I_{it}$  in (P) with  a realized stochastic shock $\varepsilon^p_{it}$ that obeys the exogeneity condition (E);  behavior variables $B_{it}$ are determined by setting $\iota = I_{it}$ and $p= P_{it}$ in  (SO) with a shock $\varepsilon^b_{it}$ that obeys (E); finally, the outcome $Y_{i, t + \ell}$ is realized by setting $\iota = I_{it}$, $p= P_{it}$, and $b = B_{it}$ in  (SO) with a shock $\varepsilon^y_{it}$ that obeys (E). \end{quote}

@@

Given the policies, we generate the counterfactual
outcomes, behavior, and information by propagating the dynamic equations:
\begin{equation} \label{eqn:CF-SEM} \tag{CF-SEM}
  \begin{aligned}
& {\ycolor Y^\star_{i,t+\ell}}  & := & Y_{i,t+\ell} ( {\bcolor B^\star_{it} } , {\pcolor P^\star_{it}}, {\icolor I^\star_{it}}), \\
& {\bcolor B^\star_{it} } & := &   B_{it}({\pcolor P^\star_{it} } ,{\icolor I^\star_{it}}), \\
& {\icolor I^\star_{it} }& := &  I_t ( {\ycolor Y^\star_{it}}, \icolor I^\star_{i,t-\ell}),  \end{aligned}
\end{equation}
with the same initialization as the factual system up to $t \leq 0$. In stating this counterfactual system of equations, we assume that structural outcome equations (SO) and information equations (I) remain invariant and so do the stochastic shocks, decomposed
into observable and unobservable parts. Formally, we record this assumption and above discussion as follows.
